---
title: "Problem Set 4"
author: "Jeff Huang"
date: "2/26/2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)

elec_resp <- read_csv(file = "ps_4_elections-poll-nc09-3.csv",
                      col_types =  cols(
                        .default = col_character(),
                        final_weight = col_double(),
                        timestamp=col_datetime(format="%m/%d/%y %H:%M")))

# Question 1.2 
rep_fav_count <- elec_resp %>% 
  filter(hrep_fav =="Favorable") %>%
  count()

und_count <- elec_resp %>%
  filter(response =="Und") %>% 
  count()

# Question 1.4
elec_resp %>% filter(response=="Dem") %>% count()

# Question 1.5
first_rep <- elec_resp %>% 
  filter(response=="Rep") %>% 
  arrange(timestamp) %>% 
  head(1)
  
  
first_dem <- elec_resp %>% 
  filter(response=="Dem") %>% 
  arrange(timestamp) %>% 
  head(1) 


```

## Question 1
There were `r elec_resp %>% filter(response=="Dem") %>% count()` respondents who supported the Democratic candidate.

There were `r rep_fav_count - und_count` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r elec_resp %>% filter(gender!=gender_combined) %>% count()` individuals for whom these variables have different values.

There are `r elec_resp %>% filter(race_eth == "White") %>% filter(file_race_black != "White") %>% count()` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

The first response of Rep came `r minute(first_rep$timestamp) - minute(first_dem$timestamp)` minutes (rounded to the nearest minute) before the first response of Dem.


```{r make_chart, echo=FALSE}
# orig %>%
#   select(response,educ4,final_weight) %>%
#   filter(response %in% c("Dem","Rep","Und"), !educ4 %in% "[DO NOT READ] Don't know/Refused") %>% 
#   group_by(educ4, response) %>% 
#   summarize(total=sum(final_weight)) %>% 
#   spread(response, total) %>% 
#   mutate(all=Dem+Rep+Und) %>% 
#   mutate(Dem=Dem/all) %>% 
#   mutate(Rep=Rep/all) %>% 
#   mutate(Und=Und/all) %>% 
#   select(-all)
```