---
title: "Problem Set #4"
author: "Jeff Huang"
date: "2/26/2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(gt)

elec_resp <- read_csv(file = "ps_4_elections-poll-nc09-3.csv",
                      col_types =  cols(
                        .default = col_character(),
                        final_weight = col_double(),
                        timestamp=col_datetime(format="%m/%d/%y %H:%M")))
# Question 1.1
dem_count <- elec_resp %>% 
  # filtering for only Democratic responses
  filter(response=="Dem") %>% 
  count()

# Question 1.2 
rep_fav_count <- elec_resp %>% 
  # filtering for only Republican responses
  filter(response =="Rep") %>%
  count()

und_count <- elec_resp %>%
  # filtering for only Undecided responses
  filter(response =="Und") %>% 
  count()

# Question 1.3
gen_diff <- elec_resp %>% 
  # comparing the two gender variables and selecting ones that do not match
  filter(gender!=gender_combined) %>% 
  count()

# Question 1.4
race_diff <- elec_resp %>% 
  #first filter for white under race_eth
  filter(race_eth == "White") %>% 
  #then filter for non_white for race_eth == white
  filter(file_race_black != "White") %>% 
  count()


# Question 1.5 - only goes to closest minute, no need to round
first_rep <- elec_resp %>% 
  # filtering for only Republican responses
  filter(response=="Rep") %>% 
  # arranging in order of timestamp and picking the first one
  arrange(timestamp) %>% 
  head(1)
  
  
first_dem <- elec_resp %>% 
  # filtering for only Democratic responses
  filter(response=="Dem") %>% 
  # arranging in order of timestamp and picking the first one
  arrange(timestamp) %>% 
  head(1) 


```

## Question 1
There were `r dem_count` respondents who supported the Democratic candidate.

There were `r rep_fav_count - und_count` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r gen_diff` individuals for whom these variables have different values.

There are `r race_diff` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

The first response of Rep came `r minute(first_rep$timestamp) - minute(first_dem$timestamp)` minutes (rounded to the nearest minute) before the first response of Dem.

## Question 2
```{r question_2, echo=FALSE}
elec_resp %>% 
  # selecting the party response, race, and final_weight to weight responses
  select(response,race_eth,final_weight) %>% 
  # filter out people who did not respond for race
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>% 
  # handles leveling to display in the right order
  mutate(race_eth=fct_relevel(race_eth, "White", "Black", "Hispanic", "Asian", "Other")) %>% 
  # groups the data by race, and then by response within those race categories
  group_by(race_eth, response) %>% 
  
  # weighting - adds up weighting of each category
  summarize(total=sum(final_weight)) %>% 
  # reorients table to display results across instead of in rows, fill=0 to handle NA
  spread(key=response, value=total, fill=0) %>% 
  # weighting each category
  mutate(all=Dem+Rep+Und+`3`) %>% 
  mutate(Dem=Dem/all) %>%
  mutate(Rep=Rep/all) %>%
  mutate(Und=Und/all) %>% 
  # filtering out unnecessary categories (-3 is a category that is not Dem, Rep, Und)
  select(-all, -"3") %>% 
  
  # special command
  ungroup() %>% 
  
  # table presentation
  gt() %>% 
    # creates title, subtitle, and source
    tab_header(
      title= "Polling Results - NC-9 District (2018)",
      subtitle="Poll results by race") %>% 
    tab_source_note("Source: The Upshot Live Poll Results (2018)") %>% 
      
    # renames the tabel column labels
    cols_label(
      race_eth = "Race",
      Dem = "DEM.",
      Rep = "REP.",
      Und = "UND."
      ) %>%
    
    # creates flat percentages
    fmt_percent(columns = vars(Dem, Rep, Und),
                decimals = 0) %>% 
  
    # creates dash instead of 0% — if column is 0, formats to display as missing
    na_if(0) %>% 
    fmt_missing(columns = vars(Und), rows = 4)

```

## Question 3
```{r question_3, echo=FALSE, fig.width = 7.2}
# figure width calibrated to display long subtitle
elec_resp %>% 
  # selecting only the education level and the weight assigned to each response 
  select(educ,final_weight) %>% 
  # filter out non-answers in the education level section
  filter(! educ %in% "[DO NOT READ] Refused") %>% 
  # arranges categories to display in the correct order through levels — reversed because the coord_flip command flips the display order
  mutate(educ=fct_rev(factor(educ, levels=c("Graduate or Professional Degree", "Bachelors' degree", "Some college or trade school", "High school", "Grade school")))) %>% 
  # groups responses by education
  group_by(educ) %>% 
  
  # graph
  ggplot(aes(x=educ,y=final_weight)) +
  # creates violin plot
  geom_violin() +
  # creates scatter plot, alpha reduced to create transparent effect, no legend
  geom_jitter(aes(alpha=0.05),position=position_jitter(0.2), show.legend = FALSE) +
  # flips the graph to display sideways
  coord_flip() +
  # x&y labels
  xlab(NULL) +
  ylab("Weight Given to Respondent in Calculating Poll Results") + 
  # titles
  labs(title = "More Educated Matter Less in North Carolina 9th Congressional District",
       subtitle = "Poll gives more weight to people who are less likely to participate in polls",
       caption="New York Times Upshot/Siena College 2018 live polls")
```

## Question 4

```{r question_4, warning=FALSE, echo=FALSE}

# I wanted to explore how polling responses differed by gender. I thought it would be interesting to do a bar graph to visualize the split between male and female percentages, and their respective quantities at first glance. I chose to represent the graph sideways to make comparisons of quantity easier on the eyes.

gen_resp <- elec_resp %>% 
  # selecting for gender and response
  select(gender, response) %>% 
  # rearranging levels to display in the right order when flipped
  mutate(response=fct_rev(factor(response, levels=c("Dem", "Rep", "Und")))) %>%
  # want the groupings of gender and responses within them
  group_by(gender, response) %>% 
  # filter to select for major party or undecided responses
  filter(response %in% c("Dem", "Rep", "Und")) %>%  
  # creates the count for each category of response and gender
  summarize(count=n())
  
  # handles graph, with fill representing the gender category 
  ggplot(gen_resp, aes(x=gen_resp$response, y=gen_resp$count, fill=gen_resp$gender, show.legend=FALSE)) +
  # creates bar graph
  geom_col() + 
  
  # flips graph 
  coord_flip() +
   
  # titles legend
  guides(fill=guide_legend(title="Gender")) +
  # major graph titles
  labs(
    title="Polling Results in North Carolina's 9th District - 2018",
    subtitle = "By Gender",
    caption = "Source: The Upshot Live Poll Results (2018)"
  ) +
  # x and y labels
  xlab("Party Response") + 
  ylab(NULL)


```

